---
title: "annotate"
author: "Huiwen Duan"
date: "10/26/2019"
output: html_document
---
```{r}
#install.packages("rfUtilities")
#install.packages("BioMed")
#install.packages("fastDummies")

library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(stringr)
library(boot)
library(randomForest)
library(rfUtilities)
library(caret)
```
read in data 
```{r warning = FALSE, message = FALSE}
anno<-read_xlsx("annotations.xlsx")
anno_orig<-anno
any(duplicated(anno_orig[,1]))
all(is.na(anno_orig$detection_id))
# get rid of detection_id column, which is empty, and the row number column
anno<-anno%>%dplyr::select(-detection_id,-`...1`)
# get rid of duplicated rows
anno<-anno%>%distinct()
```
# Exploratory Analysis 
## Explore and clean up variables:
+ Predictor variables: 
1. world_center_zyx: character, the marking in the real world coordinate system
     - plot the coordinates in 3D
     - separate it into three variables x y z
2. radius_mm: numerical, radius of a marking in mm
3. classification: categorical, annotator's classification. NAs(n=11520) means machine's annotations. so replace NAs with 999
4. user_id: categorical variable. nas(n=11520) means machine, so replace NA with 0.
     - machine and human mark on the same observations. consider random effect mixture model 
5. series_id: character, the diameter in mm of the region over which the imagery data were collected. 
     - All starts with "2.25.", so get rid of "2.25." and
     - Make it numerical 
6. val_risk: categorical, 1-4/na
7. val_classification: categorical, 1-7/na
8. data_collection_diameter: categorical with two levels
9. manufacturer: categorical, 
10. kvp: categorical with 2 levels, peak kilo voltage output of the x-ray generator used
11. manufacturer_model: categorical with 7 levels
12. recon_kernel: categorical with 9 levels, the algorithm used to build the 3D image
13. body_part: categorical with 2 levels, chest versus other. change nas to other 
14. recon_diameter: numerical variable, diameter in mm of the region from within which data were used in creating the reconstruction of the image
15. image_type: categorical with 3 levels
16. detection_id: all NAs, eliminate.
17. prob_fpr: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
18. prob_rpn: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
19. nodule_score: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 

+ human specific variables: 
3. classification: categorical, annotator's classification. NAs(n=11520) means machine's annotations. so replace NAs with 999
6. val_risk: categorical, 1-4/na
7. val_classification: categorical, 1-7/na

+ machine specific variables
17. prob_fpr: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
18. prob_rpn: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
19. nodule_score: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
      
response variable: GT, categorical, 0/1

```{r}
anno[is.na(anno$user_id),"user_id"]<-0
anno[is.na(anno$body_part),"body_part"]<-"OTHER"
anno$recon_diameter<-as.numeric(anno$recon_diameter)
anno$data_collection_diameter<-as.numeric(anno$data_collection_diameter)
row.names(anno)<-1:nrow(anno)
```
## Explore consensus on classification and val_classification
```{r}
#get positive machine GT observations
trueHumanGT<-anno_data[anno_data$GT == 1& anno_data$user_id!=0,]
mean(as.numeric(trueHumanGT$classification==trueHumanGT$val_classification))
t<-table(trueGT$classification,trueGT$val_classification)
heatmap(t, scale="column",Colv = NA, Rowv = NA,xlab="classification", ylab="val_classification")
```
## Explore world_center_zyx 
+ world_center_zyx is the unique index of nodules. An nodule may be annotated by 1-6 annotators, including machine
```{r}
#extract coordinates from world_center_zyx
coordinates<-str_extract_all(anno$world_center_zyx,"-?\\d+\\.?\\d+(e[+-]?\\d+)?",simplify = TRUE)
coordinates<-as.data.frame(coordinates)
names(coordinates)<-c("z","y","x")
coordinates$radius<-anno$radius_mm
```

```{r}
# complete anno data with coordinates
anno_data<-cbind(coordinates[,-4],anno)
anno_data<-anno_data%>%dplyr::select(-world_center_zyx)
str(anno_data)
```
```{r}
#make xyz numerical
anno_data[,1:3]<-apply(anno_data[,1:3],2,as.character)
anno_data[,1:3]<-apply(anno_data[,1:3],2,as.numeric)
```


## Explore machine versus human annotations
851 observations have both human and machine annotated 
```{r}
# subset machine and human annotations
machine_anno<-anno%>%filter(user_id==0)
human_anno<-anno%>%filter(user_id!=0)
length(unique(anno$world_center_zyx))

cross_machine<-machine_anno[which(machine_anno$world_center_zyx%in%human_anno$world_center_zyx),]
cross_human<-human_anno[which(machine_anno$world_center_zyx%in%human_anno$world_center_zyx),]

#plot GT against machine/human
is.machine<-as.numeric(anno_data1$user_id == 0)
counts <- table(is.machine,1-anno_data1$GT)
GT_mh<-barplot(counts,col=c("black","light yellow"), names.arg=c("Machine Annotations", "Human Annotations"),legend = c("Incorrect","Correct"),xlim=c(0,4))
```
## Explore radius_mm 
```{r}
anno_data1<-anno_data%>%dplyr::select(-classification,-val_risk,-val_classification,-prob_fpr,-prob_rpn,-nodule_score)
testSet<-anno_data1%>%select(-x,-y,-z)
testSet<-cbind(testSet,id[,-1])
testSet$identifier<-anno$world_center_zyx

machine_anno<-testSet[testSet$user_id==0,]
human_anno<-testSet[testSet$user_id!=0,]
testMachine<-machine_anno[which(machine_anno$identifier%in%human_anno$identifier),]
crosses<-machine_anno[which(machine_anno$identifier%in%human_anno$identifier),"identifier"]
testHuman<-human_anno[which(human_anno$identifier%in%crosses),]

# x is human, y is machine 
df<-data.frame(type = c(rep("Human",851),rep("Machine",851)),radius=c(shared$radius_mm.x,shared$radius_mm.y))
dim(shared)
shared<-merge(testHuman_distinct,testMachine,by="identifier")

df5<-data.frame(type = c(rep("Human",nrow(human_anno)),rep("Machine",nrow(machine_anno))),radius=c(human_anno$radius_mm,machine_anno$radius_mm))

#plot radius against human/machine in shared coordinates
ggplot(df, aes(x=type, y=radius)) + 
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))
#plot radius against human/machine in all data 
ggplot(df5, aes(x=type, y=radius)) + 
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))

length(unique(anno$world_center_zyx))
```

## Explore machine annotation variables
+ there are 10 observations with prob_rpn true. They include the only one true prob_fpr, the only seven true nodule_score
```{r}
summary(anno_data)
str(anno_data)
anno_data%>%filter(!is.na(prob_rpn))%>%filter(prob_rpn)
```

# Build Model to Predict GT
Predictor variables: 
1. coordinates: x,y,z numerical 
2. radius_mm: numerical, radius of a marking in mm
(3.) classification: categorical, annotator's classification. NAs(n=11520) means machine's annotations. so replace NAs with 999
4. user_id: categorical variable. nas(n=11520) means machine, so replace NA with 0.
     - machine and human mark on the same observations. consider random effect mixture model 
5. series_id: numerical, the diameter in mm of the region over which the imagery data were collected. 
(6.) val_risk: categorical, 1-4/na
(7.) val_classification: categorical, 1-7/na
8. data_collection_diameter: numerical with two levels
9. manufacturer: categorical, 
10. kvp: categorical with 2 levels, peak kilo voltage output of the x-ray generator used
11. manufacturer_model: categorical with 7 levels
12. recon_kernel: categorical with 9 levels, the algorithm used to build the 3D image
13. body_part: categorical with 2 levels, chest versus other. change nas to other 
14. recon_diameter: numerical variable, diameter in mm of the region from within which data were used in creating the reconstruction of the image
15. image_type: categorical with 3 levels
16. prob_fpr: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
17. prob_rpn: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
18. nodule_score: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
```{r}
# factorize categorical variables 
anno_data$z<-as.numeric(anno_data$z)
anno_data$y<-as.numeric(anno_data$y)
anno_data$x<-as.numeric(anno_data$x)
anno_data$classification<-as.factor(anno_data$classification)
anno_data$user_id<-as.factor(anno_data$user_id)
anno_data$GT<-as.factor(anno_data$GT)
str(anno_data[,9:21])
anno_data[,9:21]<-apply(anno_data[,9:21],2,as.factor)
```

model1: logistic regression
predictors: 
1. coordinates: x,y,z numerical 
2. radius_mm: numerical, radius of a marking in mm
(3.) classification: categorical, annotator's classification. NAs(n=11520) means machine's annotations. so replace NAs with 999
4. user_id: categorical variable. nas(n=11520) means machine, so replace NA with 0.
     - machine and human mark on the same observations. consider random effect mixture model 
5. series_id: id of a CT series. Change to dummy variables. 
(6.) val_risk: categorical, 1-4/na
(7.) val_classification: categorical, 1-7/na
8. data_collection_diameter: numerical with two levels
9. manufacturer: categorical, 
10. kvp: categorical with 2 levels, peak kilo voltage output of the x-ray generator used
11. manufacturer_model: categorical with 7 levels
12. recon_kernel: categorical with 9 levels, the algorithm used to build the 3D image
13. body_part: categorical with 2 levels, chest versus other. change nas to other 
14. recon_diameter: numerical variable, diameter in mm of the region from within which data were used in creating the reconstruction of the image
15. image_type: categorical with 3 levels
(16.) prob_fpr: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
(17.) prob_rpn: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
(18.) nodule_score: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 

Excluding human specific and machine specific variables, which are
+ human specific variables: 
3. classification: categorical, annotator's classification. NAs(n=11520) means machine's annotations. so replace NAs with 999
6. val_risk: categorical, 1-4/na
7. val_classification: categorical, 1-7/na

+ machine specific variables
17. prob_fpr: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
18. prob_rpn: TRUE/FALSE/NA(human annotators)
      - change to three level categorical 
19. nodule_score: TRUE/FALSE/NA(human annotators)
      - change to three level categorical

## Excluding human specific and machine specific variables
### Final Model: random forest  
```{r}
anno_data1<-anno_data%>%dplyr::select(-classification,-val_risk,-val_classification,-prob_fpr,-prob_rpn,-nodule_score)
str(anno_data)
for(i in c(9:13,15)){
  anno_data1[,i]<-as.factor(anno_data1[,i])
}
```

```{r} 
set.seed(1998)
perf<-list()
# test out hyperparameters
#for (i in 5:10){
#  modelRF<-randomForest(GT~.,data=anno_data1,mtry = i)
#  predictionRF<-modelRF$predicted
#  perf[[i-4]]<-confusionMatrix(anno_data1$GT,modelRF$predicted,positive = "1")
#}
anno_data1$identifier<-anno$world_center_zyx
anno_data1_distinct<-anno_data1[!duplicated(anno_data1$identifier),]%>%dplyr::select(-identifier)
```
explore series_id variable
```{r}
#make dummy variables
id<-fastDummies::dummy_cols(as.factor(as.character(anno_data$series_id)))
names(id)[2:length(id)]<-paste("V", 1:(ncol(id)-1), sep = "_")
id_GT<-cbind(anno_data$GT,id[,-1])
id_GT<-data.frame(id_GT)
```
inplement random forest model
```{r}
test_data3<-anno_data1%>%select(-series_id,-x,-y,-z)
test_data3<-cbind(test_data3,id[,-1])
test_data3$identifier<-anno$world_center_zyx
test_data3<-test_data3[!duplicated(test_data3$identifier),]
test_data3<-test_data3%>%select(-identifier)
table(test_data3$manufacturer)
t3<-randomForest(factor(GT)~.,data=test_data3,mtry = 5,importance=TRUE)
predictiont3<-t3$predicted
confusionMatrix(factor(test_data3$GT),predictiont3,positive = "1")
# get variable importance
varImpPlot(t3,n.var = 10)
str(anno_data1)
```
test the model on human/machine shared coordinates
```{r}
testSet<-anno_data1%>%select(-series_id,-x,-y,-z)
testSet<-cbind(testSet,id[,-1])
testSet$identifier<-anno$world_center_zyx
#machine annotations
machine_anno<-testSet[testSet$user_id==0,]
#human annotations
human_anno<-testSet[testSet$user_id!=0,]

#machine annotations on the shared coordinates
testMachine<-machine_anno[which(machine_anno$identifier%in%human_anno$identifier),]
crosses<-machine_anno[which(machine_anno$identifier%in%human_anno$identifier),"identifier"]
#human annotations on the shared coordinates
testHuman<-human_anno[which(human_anno$identifier%in%crosses),]
testMachine<-testMachine%>%dplyr::select(-identifier)
testHuman<-testHuman%>%dplyr::select(-identifier)

p1<-predict(t3,testMachine)
confusionMatrix(factor(testMachine$GT),p1,positive = "1")

p2<-predict(t3,testHuman)
confusionMatrix(factor(testHuman$GT),p2,positive = "1")
```
## Exploratory Modelings
### Part2: human only
```{r}
#exclude human specific and machine specific variables
humanRF<-randomForest(GT~.,data=anno_data1%>%filter(user_id!=0),mtry = 5)
human_data1<-anno_data1%>%filter(user_id!=0)
confusionMatrix(human_data1$GT,humanRF$predicted,positive = "1")
```

```{r}
#include human specific variables
human_data2<-anno_data%>%dplyr::select(-prob_fpr,-prob_rpn,-nodule_score)%>%filter(user_id!=0)
human_data2[is.na(human_data2$val_classification),"val_classification"]<-99
human_data2[is.na(human_data2$val_risk),"val_risk"]<-99
dim(human_data2)
for(i in c(5:16,18)){
  human_data2[,i]<-as.factor(human_data2[,i])
}

summary(human_data2)
human_data2<-human_data2%>%dplyr::select(-series_id)

human_data2$body_part<-as.numeric(human_data2$body_part)

humanRF2<-randomForest(GT~.,data=human_data2,mtry = 5)
humanRF2$importance
t<-as.factor(human_data2$val_classification)
table(t)
```

### Part3: machine only
```{r}
machine_data1<-anno_data1%>%filter(user_id==0)
machineRF<-randomForest(GT~.,data=machine_data1,mtry = 5)
confusionMatrix(machine_data1$GT,machineRF$predicted,positive = "1")
table(machine_data1$GT)
```
```{r}
index<-sample(nrow(anno_data1),0.9*nrow(anno_data1))
training<-anno_data1[index,]
testing<-anno_data1[-index,]
model1<-glm(GT~.,data = training,family="binomial")
#summary(model1)
prediction<-predict(model1,newdata = testing,type="response")
?predict
prediction_class<-as.numeric(prediction>0.5)
mean(prediction_class == testing$GT)

```

# Interactive plots of coordinates
```{r}
# plot coordinates in the same series_id in machine annotations
machine_anno<-anno_data[anno_data$user_id==0,]
frequent_series<-table(machine_anno$series_id)%>%sort(decreasing = TRUE)%>%head(5)
anno_data$series_id<-as.factor(anno_data$series_id)

df1<-machine_anno[which(machine_anno$series_id==names(frequent_series[1])),]

dim(df1)
p <- plot_ly(df1, x = ~x, y = ~y, z = ~z,marker = list(color = ~radius_mm, colorscale = c('#FFE1A1', '#683531'), showscale = TRUE, main = "machine")) %>%
  add_markers() %>%
  layout(title = "Machine Annotations Under the Most Frequent Series_id",
         scene = list(xaxis = list(title = 'x'),
                     yaxis = list(title = 'y'),
                     zaxis = list(title = 'z')))
p

```

```{r}
# plot coordinates in the same series_id with both human and machine annotations
frequent_series_c<-table(anno_data$series_id)%>%sort(decreasing = TRUE)%>%head(5)
anno_data$series_id<-as.factor(anno_data$series_id)

dff1<-anno_data[which(anno_data$series_id==names(frequent_series_c[1])),]
dff1$is_machine<-as.numeric(dff1$user_id==0)
dim(dff1)
p <- plot_ly(dff1, x = ~x, y = ~y, z = ~z,marker = list(color = ~is_machine, colorscale = c('white', 'red'))) %>%
  add_markers() %>%
  layout(title = "Machine&Human Annotations Under the Most Frequent Series_id"
         ,scene = list(xaxis = list(title = 'x'),
                     yaxis = list(title = 'y'),
                     zaxis = list(title = 'z')))
p
```

